### 2022-01-27 thesis.tex

% !BIB TS-program = biber
\documentclass[onehalf,11pt]{beavtex/beavtex}

\input{thesis-header.tex}

% things to check or replace
% replace your name here.
\newcommand{\TheAuthors}{Derek Bean}
\title{Ignition of Fuel Beds by Firebrands}
\degree{Doctor of Philosophy}
\major{Mechanical Engineering}

\submitdate{June 13, 2018}
\commencementyear{2019}

\author{\TheAuthors{}}
\doctype{Thesis}
\department{Mechanical, Industrial, and Manufacturing Engineering}
\depttype{School}
\depthead{Head}

\advisor{David Blunck}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% figure location
\graphicspath{{./Figures/}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\abstract
{
(ABSTRACT GOES HERE)
}

\acknowledgements{
\textit{here you might thank the people that helped you, and possibly a funding source}
}

\dedication{(optional dedication)}

\begin{document}

\maketitle

\mainmatter

% By including `''%!TEX root = thesis.tex` in the top of each included file,
% most LaTeX editors will automatically compile the main package when you are working
% within a file.

\include{introduction}
\include{litreview}
\include{manuscript1}
\include{manuscript2}
\include{manuscript3}
\include{results}
\include{conclusions}

\newpage

\printbibliography[]

%% INCLUDE IF YOU NEED AN APPENDIX

% \newpage
%\appendix
%\input{appendixA.tex}

\end{document}