### 2022-05-25 thesis.tex

% !BIB TS-program = biber
\documentclass[onehalf,11pt]{beavtex/beavtex}

\input{thesis-header.tex}

% things to check or replace
% replace your name here.
\newcommand{\TheAuthors}{Derek Bean}
\title{Ignition of Fuel Beds by Firebrands}
\degree{Doctor of Philosophy}
\major{Mechanical Engineering}

\submitdate{June 8, 2022}
\commencementyear{2022}

\author{\TheAuthors{}}
\doctype{Thesis}
\department{Mechanical, Industrial, and Manufacturing Engineering}
\depttype{School}
\depthead{Head}

\advisor{David Blunck}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% figure location
\graphicspath{{./Figures/}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\abstract
{
(ABSTRACT GOES HERE)
}

\acknowledgements{
I have been fortunate to be perpetually surrounded by people that have given me an immense amount of support and guidance throughout my life and this has been especially true in my studies. This period of my life has been truly transformative and it would not have been possible without all of the wonderful people in my life. I owe the greatest thanks to my family whose unwavering support, encouragement, and guidance has been invaluable. I would not be who I am or where I am today if it was not for you all. It is impossible to put into words how much I appreciate everything you have done for me. To Heather, thank you for being my partner through the toughest parts of this journey, for asking so many great questions, and encouraging me to keep going so that I may reach my goals. I doubt I would have made it this far without you.

Thank you to Dr. David Blunck for giving me the opportunity to conduct this work. In the graduate school experience you have facilitated I have achieved things I did not know I was capable of. I have the deepest gratitude for your guidance, insight, and the example that you set. I owe a similar debt of gratitude to my fellow students....

This work was supported by the National Institute of Standards and Technology as part of project number 70NANB17H281.
}

%\dedication{(optional dedication)}

\begin{document}

\maketitle

\mainmatter

% By including `''%!TEX root = thesis.tex` in the top of each included file,
% most LaTeX editors will automatically compile the main package when you are working
% within a file.
%\printnomenclature
\include{introduction}
\include{litreview}
\include{manuscript1}
\include{manuscript2}
\include{manuscript3}
\include{manuscript4}
%\include{results}
\include{conclusions}

\newpage

\printbibliography[]

%% INCLUDE IF YOU NEED AN APPENDIX

% \newpage
%\appendix
%\input{appendixA.tex}

\end{document}